---
title: "BrewData Codebook ** Top Secret **"
author: "Jim Park & Randall Hendrickson"
date: "October 13, 2017"
output: 
   html_document:
      keep_md: true
editor_options: 
  chunk_output_type: console
---

To access data and code for this codebook, please fork us at:
https://github.com/rhendrickson42/msds6306Case

# The following section is the codebook for the Master_JP_RH.Rmd document.

PACKAGES/LIBRARIES
devtools::install_github("krlmlr/here") - Used to simplify working directory.  Self determines.

library('plyr') # to call the count command which provies a frequency of occurance for a factor.  Used in this case to count occurances of breweries in each state.

library(here) - Used to simplify working directory.  Self determines. Reads CSV files from data file and returns a local file reference.

library(ggplot2) - a plotting package used for several visuals in this report.


OBJECTS/DATA FILES
Brewies.csv - Contians brewery information including Brew_ID,Name,City, and State

Beers.csv - Contians Beer information including Name, Beer_ID, ABV, IBU, Brewery_id and Style,Ounces for 2410 observations

Breweries - data frame object with brewery information created by reading Breweries.csv

Beers - data frame object with beers information 

States - an object containing a list of states.  Each entry represents one brewery in that state later used for number of breweries (frequency) in the state.

mergedData - a data frame which consists of two merged objects :Beers and Breweries.  Renamed columns and tidy (cleared up white space.  Euro characters were left as is and had no impact) where necessary

num_NAs - an object that consists of a summary of NA's within each column of the mergedData dataframe.

state_ABV- an object that consists of the median ABV statistics by state extracted from the mergedData data frame.

State_IBU - an object that consists of the median IBU statistics by state extracted from the mergedData data frame.

StateQty - a data frame that aggregates the state name and the number of times it occurs in the mergedData data frame.  This data frame is built in a loop.

StateBeers - an object with a subset of information from mergedData data frame that consists of the selected state and all of the beers in that state.

ColorOfPlot - an object used to set the aesthetics of the ABV plot

ColorOfPlotIBU - an object used to set the aesthetics of the IBU plot

BeersInStateOutput - an object used to construct a state specific output sentance.

StateCenter - a dataframe used to list the state, Latitude, and Longitude.

StateLatLong - a dataframe used to list the state, Latitude, and Longitude including DC as a state.

StateCenterABV - a dataframe consisting of a merge of StateLatLong and ABV data from StateCenter

sa - an object with state abbreviations (from a built in package called state.abb)

cs - an object with state center latitude and longitude (from a built in package called state.center)

# The following section is the codebook for customer specific reports in the "markdown" directory
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
if(!require("here")) cat("please install here package # devtools::install_github(\"krlmlr/here\")")
library(knitr)
library(here)
dataSetup_script <- here("scripts", "dataSetup.R")
analysis_script <- here("scripts", "analysis.R")
source(dataSetup_script)
```

## BrewData Codebook - Classified
This codebook outlines the **"secret sauce"** for BrewData's technical apparatus.

read chunk = dataSetup_script.R
```{r setup_codebook}
read_chunk(paste(dataSetup_script))
```

### 1. Script - **dataSetup.R**

uses the @ knitr variable pattern to read the following

- scriptDataSetup_files
- scriptDataSetup_readFiles
- scriptDataSetup_tidyBeersAndBreweries
- scriptDataSetup_logsessionInfo

#### 1. a. Locate the raw data files.
```{r scriptDataSetup_files, echo=TRUE}
```
#### Result: 
```{r echo=TRUE}
# the raw data files:
sprintf("This returns two files: %s , %s", local_file1, local_file2)
```

#### 1. b. Read the data files
```{r scriptDataSetup_readFiles, echo=TRUE}
```
#### Result: 
```{r echo=TRUE}
# This returns two data frames, beers and breweries
# structure of beers
str(script_beers)
colnames(script_beers)
# structure of breweries
str(script_breweries)
colnames(script_breweries)
```

#### 1. c. tidy the data
```{r scriptDataSetup_tidyBeersAndBreweries, echo=TRUE}
```
#### Result: tidy data

#### 1. d. Log sessionInfo
```{r scriptDataSetup_logsessionInfo, echo=FALSE}
```


### 2. Script - **analysis.R**

uses the @ knitr variable pattern to read the following

- script_Analysis_Question1
- script_Analysis_Question2
- script_Analysis_Question3
- script_Analysis_Question4
- script_Analysis_Question5
- script_Analysis_Question6
- script_Analysis_Question7


read chunk = analysis.R
```{r setup_codebook2}
read_chunk(paste(analysis_script))
#source(analysis_script)
```

Questions:

#### 2.1 How many breweries are present in each state?
```{r script_Analysis_Question1, echo=TRUE}
```
#### Result: 
```{r echo=TRUE}
number_breweries_byState
```

#### 2.2 Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
```{r script_Analysis_Question2, echo=TRUE}
```
#### Result: 
```{r echo=TRUE}
# print first 6 observations
head(script_beer_world, 6)

# print last 6 observations
tail(script_beer_world, 6)
```

#### 2.3 Report the number of NA's in each column.
```{r script_Analysis_Question3, echo=TRUE}
```
#### Result: 
```{r echo=TRUE}
# report number of NAs
num_NAs
```

#### 2.4 Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare. 
```{r script_Analysis_Question4, echo=TRUE}
```
#### Result: 
```{r echo=TRUE}

```

#### 2.5 Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r script_Analysis_Question5, echo=TRUE}
```
#### Result: 
```{r echo=TRUE}
# max ABV
script_state_max_ABV

# max ABV
script_state_max_IBU
```

#### 2.6 Summary statistics for the ABV variable.
```{r script_Analysis_Question6, echo=TRUE}
```
#### Result: 
```{r echo=TRUE}
# printed summary of ABV
```

#### 2.7 Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot. Make your best judgment of a relationship and EXPLAIN your answer.
```{r script_Analysis_Question7, echo=TRUE}
```
#### Result: 
```{r echo=TRUE}
# Summary of relationship
```

To access data and code for this codebook, please fork us at:
https://github.com/rhendrickson42/msds6306Case




