---
title: "Case Study 01 - Beers"
author: "Randall Hendrickson and Jim Park"
date: "October 6, 2017"
params:
  state: "TX"
output: 
   html_document:
      keep_md: true
---

#INTRODUCTION

The number of breweries in the United States has more than tripled in the last decade according to www.brewersassociation.org.  Along with this explosive growth the industry is becoming much more competitive.   For startup breweries it is important to understand the level of competition within the state and the preferences of the consumer in order to be successful.  BrewData Inc., a data science company specializing in brewery and beer data, is pleased to provide our client, Great-Beer Distributors Inc. (GBDI), with an easy to use R program which reports out specific beer and brewery data of interested by the state as well as national trends.  As requested BrewData Inc. will conduct this work utilizing the data provided by GBDI.  Finally, GBDI has also asked BrewDAta Inc. to simplify the data provided by consolidating in to one merged data file and to identify missing data.

CAPABILITIES AND DEMONSTRATION OF PROGRAM
To illustrate the program capabilites BrewData Inc. has provided sample program output using the state of Texas. To get additional reports for different states, change the state in *line 6* of the *Master_JP_RH.Rmd* document to another state, and run the *make* command at the project root directory, or use RStudio's *Build All* button. The program will provide a report specific to the state entered.  A sample report for the state of Texas, entered as 'TX', is provided below.

Here are specific responses as requested by our contract signed earlier. Please note, for the purpose of this report,  Washington D.C is modeled as a separate state.

1. As requested by GBDI, we have the following data for the number of breweries in each state. This will provide data whether their new brewery has local competition and by how many breweries.

```{r, ECHO = TRUE}
StateSelected <- params$state
# devtools::install_github("krlmlr/here")
library(here)
local_file1 <- here("data", "Breweries.csv")
local_file2 <- here("data", "Beers.csv")
Breweries <- read.csv(paste(local_file1),sep=",", header = TRUE)
library('plyr') # to call the count command
States <- (Breweries$State) #assigns variable to States
count(States) #Generates frequency table.  Number of Observations(breweries)/State
```

2. Additional information requested by GBDI was related to the beers that these respective brewers produce. We have consolidated this data by merging the data for the beers produced by each brewer. As requested, here are the first and last six observations for this data. A printing of the first 6 observations and the last 6 observations are provided here as a sample of the output.

```{r Q2MergedData, ECHO=TRUE}
Beers <- read.csv(paste(local_file2),sep=",", header = TRUE)
mergedData <- merge(Breweries, Beers, by.x=c('Brew_ID'), by.y=c('Brewery_id'))
names(mergedData)[names(mergedData) == "Name.x"] <- "Brewery_Name"
names(mergedData)[names(mergedData) == "Name.y"] <- "Beer_Name"
mergedData$State <- gsub(" ", "", mergedData$State, fixed = TRUE) #Removes extra spaces that area not needed.
head(mergedData,n=6)
tail(mergedData, n=6)
```

3. As requested by GBDI, we are pleased to perform robustness tests on the produced data. We have identified some missing data as detailed below. This is represented by an "NA" term and the counts are detailed below. The chart below shows there are 62 NAs in Alcohol by Volume (ABV) and there are 1005 NAs in International Bitterness Unit data. Both were from data sets provided by GBDI. It should be noted that South Dakota did not have any reported IBU data. (For an additional fee, we can provide testing for all missing NA data or provide state of the art bootstrapping techniques.)

```{r question3}
# report number of NAs
num_NAs <- sapply(mergedData, function(x) sum(is.na(x)))
num_NAs
```

4. After speaking with GBDI's *Brewmaster Mr. Potterhead*, we are well aware of the need to understand the ABV and IBU medians from the provided data. We are pleased to provide *Mr. Potterhead* with the following results and he should note that the report is customized to highlight The State of Texas in *Blue* as defined in our contract.

```{r ABVIBUCharts, ECHO=TRUE}
state_ABV <- aggregate(mergedData["ABV"], by=mergedData[c("State")], FUN=median, na.rm=TRUE)
State_IBU <- aggregate(mergedData["IBU"], by=mergedData[c("State")], FUN=median, na.rm=TRUE)
library('plyr')
ListOfStates <- sort(unique(mergedData$State)) # Creates a list of the states in alphabetical order assuming it may be needed by the application user.
StateQty <- data.frame(matrix(ncol = 2, nrow = 51)) #Initiates data frame for loop
ColorOfPlot <- 0 #Initiates vector for loop to be used to color code bar chart.
library(data.table)
setnames(StateQty, old = c('X1','X2'), new = c('State','Number')) #Changes the names of columns.

#  Loop below builds a data frame with state, number of beers in each state, and color scheme for bar chart
for(i in 1:51){
StateQty$State[i] <- as.character(ListOfStates[i]) #Adds each state to the vector
StateBeers <- mergedData[grep(ListOfStates[i], mergedData$State),]
StateQty$Number[i] <- nrow(StateBeers)
ColorOfPlot[i] <- 'Red' #Initializes all to the color red.
}

StateQty <- StateQty[order(-StateQty$Number),]#Orders from high to low based on number in brewery
#StateQty
ColorOfPlot[which(grepl(StateSelected, StateQty$State))] <- 'blue' #Changes the color of the state of interest to stand out in plotting.  Feeds into fill command in plot that follows.
# Plots a histogram in decreasing order of beers per state with the state of interest highlighted
library(ggplot2)
ggplot(data=StateQty, aes(x=reorder(State, -Number), y=Number)) + geom_bar(stat = "identity", fill=ColorOfPlot) +  theme(axis.text.x=element_text(angle=90,vjust=0.5)) + xlab("State") + ylab("Number of Beers") + ggtitle("Number of Different Beers by State")

# Loop to construct the color scheme for fill option in plotting median IBU by state
ColorOfPlotIBU <- 0
for(i in 1:51){
ColorOfPlotIBU[i] <- 'Red'
}
State_IBU <- State_IBU[order(-State_IBU$IBU),]#Orders from high to low based on IBU
State_IBU[is.na(State_IBU)] <- 0
ColorOfPlotIBU[which(grepl(StateSelected, State_IBU$State))] <- 'blue' 
ggplot(data=State_IBU, aes(x=reorder(State, -IBU), y=IBU)) + geom_bar(stat = "identity", fill=ColorOfPlotIBU) +  theme(axis.text.x=element_text(angle=90,vjust=0.5)) + xlab("State") + ylab("IBU") + ggtitle("IBU Median by State") + labs(caption = "(*SD did not have any reported IBU values)")
```

Additionally, after deep analysis, we are able to provide which state has the maximum alcoholic (ABV) beer and which state has the most bitter (IBU) beer. Please find the following data for your reference.

The state with the highest (max) median ABV is District of Columbia (DC).
The state with the highest (max) median IBU is Maine (ME).

```{r question_5, ECHO=TRUE}

# max ABV
state_ABV[which.max(state_ABV$ABV),]

# max IBU
State_IBU[which.max(State_IBU$IBU),]

```

6. Following the last sections of our contract, we are providing a summary of the ABV data, this summary includes the data from the output below.

```{r question_6, ECHO=TRUE}

summary(state_ABV)

```

7. Finally, GBDi requested us to understand the relationship between the bitterness of the beer and its alcoholic content? Please see the below scatter plot. You will see there is a somewhat linear trend whereas IBU increases, so does ABV. Please note, that this is an observational study so causation cannot be inferred.

```{r beer_plot}

ggplot(mergedData, aes(x = IBU, y = ABV)) + 
  geom_point(colour = 'red', size = 1, na.rm=TRUE) + 
  labs(title = "Relationship between Bitterness and Alcohol content", subtitle = "Is bitter better?") + 
  labs(caption = "(based on data from GBDI") +
  geom_smooth(method=lm, se=FALSE, size = 1, na.rm=TRUE)

```

In addition to the information requested under contract BrewData Inc. strives to exceed our customers expectations.  We hope you find the following information useful in understanding beer adn brewery trends in the selected state and in the nation in general.  The queries the data provided and lists all of the beers in a particular state (in this case Texas) along with a custom statement.  This data may be useful in understanding the level of competition in your state as well as an indication of which states are "Brew Friendly."  Please note that the data below is simply the first ten beers and only the last 5 columns due to space considerations.

```{r BeersInState, ECHO=TRUE}
StateBeers <- mergedData[grep(StateSelected, mergedData$State),]
BeersInStateOutput <- paste('There are', length(ListOfStates), ' Unique states in this data set and', nrow(StateBeers), 'unique beers in the state of', StateSelected)
BeersInStateOutput
head(StateBeers[,5:10], n=10)
```

Finally, BrewData Inc. has provided sample output for our patent pending regional ABV preference algorithm.   This tool assesses if there is a correlation between ABV and longitude and latitude as measured by the center of each state.   In both cases the correlation was relatively weak (less than 13%) with p values of 0.5737 (Longitude) and 0.3861 (Latitude).   Althought the correlations are week, With additional contract work we could explore similar trends that your prime competitor seems to be leveraging fairly well.

# A Look at possible correlation between longitude and ABV
```{r LatLong, ECHO=TRUE}
StateLatLong <- data.frame(matrix(ncol = 3, nrow = 50))
sa <- as.data.frame(state.abb)
names(sa)[names(sa) == "state.abb"] <- "State"
sc <- as.data.frame(state.center)
StateLatLong <- cbind(sa,sc)
names(StateLatLong)[names(StateLatLong) == "x"] <- "Longitude"
names(StateLatLong)[names(StateLatLong) == "y"] <- "Latitude"
StateLatLong$State <- as.character(as.factor(StateLatLong$State))

state_ABV$State <- as.character(as.factor(state_ABV$State))
StateLatLong <- rbind(StateLatLong, c("DC", -77.0369, 38.9072)) #Adds DC to the list
StateLatLong$Longitude <- as.numeric(as.character(StateLatLong$Longitude))
StateLatLong$Latitude <- as.numeric(as.character(StateLatLong$Latitude))
StateLatLong <- StateLatLong[order(StateLatLong$State),]

StateCenterABV <-data.frame(matrix(ncol = 4, nrow = 51))
#StateCenterABV <- data.frame(matrix(ncol=4, nrow=51)
StateCenterABV <- merge(StateLatLong, state_ABV, by=c('State'))
StateCenterABV
plot(StateCenterABV$Longitude, StateCenterABV$ABV)
abline(lm(StateCenterABV$ABV~StateCenterABV$Longitude),col="blue",lwd=2)
cor.test(StateCenterABV$Longitude, StateCenterABV$ABV)
```

# A Look at possible correlation between latitude and ABV
```{r LatLong1, ECHO=TRUE}
StateLatLong <- data.frame(matrix(ncol = 3, nrow = 50))
sa <- as.data.frame(state.abb)
names(sa)[names(sa) == "state.abb"] <- "State"
sc <- as.data.frame(state.center)
StateLatLong <- cbind(sa,sc)
names(StateLatLong)[names(StateLatLong) == "x"] <- "Longitude"
names(StateLatLong)[names(StateLatLong) == "y"] <- "Latitude"
StateLatLong$State <- as.character(as.factor(StateLatLong$State))
state_ABV$State <- as.character(as.factor(state_ABV$State))
StateLatLong <- rbind(StateLatLong, c("DC", -77.0369, 38.9072)) #Adds DC to the list
StateLatLong$Longitude <- as.numeric(as.character(StateLatLong$Longitude))
StateLatLong$Latitude <- as.numeric(as.character(StateLatLong$Latitude))
StateLatLong <- StateLatLong[order(StateLatLong$State),]
StateCenterABV <-data.frame(matrix(ncol = 4, nrow = 51))
#StateCenterABV <- data.frame(matrix(ncol=4, nrow=51)
StateCenterABV <- merge(StateLatLong, state_ABV, by=c('State'))
StateCenterABV

cor.test(StateCenterABV$Latitude, StateCenterABV$ABV)
lm(StateCenterABV$ABV~StateCenterABV$Latitude) #in the form Y=X
myfit=lm(StateCenterABV$ABV~StateCenterABV$Latitude) #in the form Y=X
myfit
plot(StateCenterABV$Latitude, StateCenterABV$ABV)
abline(lm(StateCenterABV$ABV~StateCenterABV$Latitude),col="blue",lwd=2)
```

# CONCLUSION
BrewData Inc. has concluded that the beer industry is in a high growth mode in both ABV and IBU.   We would be please to collect and leverage further data in defining a profitable niche for your next super beer.

## Cheers, R. Hendrickson (Chief Beer Officer) and J. Park (Chief Data Officer) **




