---
title: "Case Study 01 - Beers"
author: "Randall Hendrickson and Jim Park"
date: "October 6, 2017"
params:
  state: "TX"
output: 
   html_document:
      keep_md: true
---

#INTRODUCTION

The number of breweries in the United States has more than tripled in the last decade according to www.brewersassociation.org.  Along with this explosive growth the industry is becoming much more competitive.   For startup breweries it is important to understand the level of competition within the state and the preferences of the consumer in order to be successful.  BrewData Inc., a data science company specializing in brewery and beer data, is pleased to provide our client, Great-Beer Distributors Inc. (GBDI), with an easy to use R program which reports out specific beer and brewery data of interested by the state as well as national trends.  As requested BrewData Inc. will conduct this work utilizing the data provided by GBDI.  Finally, GBDI has also asked BrewDAta Inc. to simplify the data provided by consolidating in to one merged data file and to identify missing data.

CAPABILITIES AND DEMONSTRATION OF PROGRAM
To illustrate the program capabilites BrewData Inc. has provided the following sample program output.   All all that is needed is to enter the state in the indicated location and the program will provide a report specific to the state entered.  A sample report for the state of Michigan, entered as 'MI', is provided below.

TODO:File reference still required in bibliography.
TODO:  Clear explicit answers to all questions.


Once the state is entered the program will output the following useful information:


1. Number of Breweries in each state.  This provides national context.
```{r, ECHO = FALSE}
StateSelected <- params$state
#tateSelected <- c('MI') #Fill in desired state inbetween the quotes here from list of States by alphabetical order.
# devtools::install_github("krlmlr/here")
library(here)
local_file1 <- here("data", "Breweries.csv")
local_file2 <- here("data", "Beers.csv")
Breweries <- read.csv(paste(local_file1),sep=",", header = TRUE)
library('plyr') # to call the count command
States <- (Breweries$State) #assigns variable to States
names(States)[names(States) == "x"] <- "State"
names(States)[names(States) == "freq"] <- "frequency"
count(States) #Generates frequency table.  Number of Observations(breweries)/State
```


2. Merge beer data with Breweries data and consolidate into one document. A printing of the first 6 observations and the last 6 observations are provided here as a sample of the output.
```{r Q2MergedData, ECHO=TRUE}
Beers <- read.csv(paste(local_file2),sep=",", header = TRUE)
mergedData <- merge(Breweries, Beers, by.x=c('Brew_ID'), by.y=c('Brewery_id'))
names(mergedData)[names(mergedData) == "Name.x"] <- "Brewery_Name"
names(mergedData)[names(mergedData) == "Name.y"] <- "Beer_Name"
mergedData$State <- gsub(" ", "", mergedData$State, fixed = TRUE) #Removes extra spaces that area not needed.
head(mergedData,n=6)
tail(mergedData, n=6)
```


3. Identification of missing data.  In this output missing data will be represented by an "NA" term.

```{r question3}
# report number of NAs
num_NAs <- sapply(mergedData, function(x) sum(is.na(x)))
num_NAs
```


4. Compute the median Alcohol By Volume content (ABV) and the International Bitterness Unite (IBU) for each state.


```{r ABVIBUCharts, ECHO=TRUE}
state_ABV <- aggregate(mergedData["ABV"], by=mergedData[c("State")], FUN=median, na.rm=TRUE)
State_IBU <- aggregate(mergedData["IBU"], by=mergedData[c("State")], FUN=median, na.rm=TRUE)
library('plyr')
ListOfStates <- sort(unique(mergedData$State)) # Creates a list of the states in alphabetical order assuming it may be needed by the application user.
StateQty <- data.frame(matrix(ncol = 2, nrow = 51)) #Initiates data frame for loop
ColorOfPlot <- 0 #Initiates vector for loop to be used to color code bar chart.
library(data.table)
setnames(StateQty, old = c('X1','X2'), new = c('State','Number')) #Changes the names of columns.

#  Loop below builds a data frame with state, number of beers in each state, and color scheme for bar chart
for(i in 1:51){
StateQty$State[i] <- as.character(ListOfStates[i]) #Adds each state to the vector
StateBeers <- mergedData[grep(ListOfStates[i], mergedData$State),]
StateQty$Number[i] <- nrow(StateBeers)
ColorOfPlot[i] <- 'Red' #Initializes all to the color red.
}

StateQty <- StateQty[order(-StateQty$Number),]#Orders from high to low based on number in brewery
#StateQty
ColorOfPlot[which(grepl(StateSelected, StateQty$State))] <- 'blue' #Changes the color of the state of interest to stand out in plotting.  Feeds into fill command in plot that follows.
# Plots a histogram in decreasing order of beers per state with the state of interest highlighted
library(ggplot2)
ggplot(data=StateQty, aes(x=reorder(State, -Number), y=Number)) + geom_bar(stat = "identity", fill=ColorOfPlot) +  theme(axis.text.x=element_text(angle=90,vjust=0.5)) + xlab("State") + ylab("Number of Beers") + ggtitle("Number of Different Beers by State")

# Loop to construct the color scheme for fill option in plotting median IBU by state
ColorOfPlotIBU <- 0
for(i in 1:51){
ColorOfPlotIBU[i] <- 'Red'
}
State_IBU <- State_IBU[order(-State_IBU$IBU),]#Orders from high to low based on IBU
State_IBU[is.na(State_IBU)] <- 0
ColorOfPlotIBU[which(grepl(StateSelected, State_IBU$State))] <- 'blue' 
ggplot(data=State_IBU, aes(x=reorder(State, -IBU), y=IBU)) + geom_bar(stat = "identity", fill=ColorOfPlotIBU) +  theme(axis.text.x=element_text(angle=90,vjust=0.5)) + xlab("State") + ylab("IBU") + ggtitle("IBU Median by State") + labs(caption = "(*SD did not have any reported IBU values)")
```


5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
The state with the highest (max) median ABV is District of Columbia (DC).
The state with the highest (max) median IBU is Maine (ME).

```{r question_5, ECHO=TRUE}

# max ABV
state_ABV[which.max(state_ABV$ABV),]

# max IBU
State_IBU[which.max(State_IBU$IBU),]

```
6. Summary statistics for the ABV variable.  TODO: Add description of how this fits the report.

```{r question_6, ECHO=TRUE}

summary(state_ABV)

```

7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot. TODO:How do we want to describe this trend?

```{r beer_plot}

ggplot(mergedData, aes(x = IBU, y = ABV)) + 
  geom_point(colour = 'red', size = 1, na.rm=TRUE) + 
  labs(title = "Relationship between Bitterness and Alcohol content", subtitle = "Is bitter better?") + 
  labs(caption = "(based on data from ...?? where's the data from ??)") +
  geom_smooth(method=lm, se=FALSE, size = 1, na.rm=TRUE)

```


8.TODO:  Add in if we have time. Number of different beers in each state highlighting how the selected state compares.  This provides insight into the level of competition in your state as well as an indication which states are "Brew Friendly." 


9.TODO:  Add in if we have time. Number of beers in your state by style.  Are there too many Pilsners already in your state? 

10. ABV preference by Latitude and Longitude of centroid of State.

# A Look at possible correlation between longitude and ABV
```{r LatLong, ECHO=TRUE}
StateLatLong <- data.frame(matrix(ncol = 3, nrow = 50))
sa <- as.data.frame(state.abb)
names(sa)[names(sa) == "state.abb"] <- "State"
sc <- as.data.frame(state.center)
StateLatLong <- cbind(sa,sc)
names(StateLatLong)[names(StateLatLong) == "x"] <- "Longitude"
names(StateLatLong)[names(StateLatLong) == "y"] <- "Latitude"
StateLatLong$State <- as.character(as.factor(StateLatLong$State))

state_ABV$State <- as.character(as.factor(state_ABV$State))
StateLatLong <- rbind(StateLatLong, c("DC", -77.0369, 38.9072)) #Adds DC to the list
StateLatLong$Longitude <- as.numeric(as.character(StateLatLong$Longitude))
StateLatLong$Latitude <- as.numeric(as.character(StateLatLong$Latitude))
StateLatLong <- StateLatLong[order(StateLatLong$State),]

StateCenterABV <-data.frame(matrix(ncol = 4, nrow = 51))
#StateCenterABV <- data.frame(matrix(ncol=4, nrow=51)
StateCenterABV <- merge(StateLatLong, state_ABV, by=c('State'))
StateCenterABV
plot(StateCenterABV$Longitude, StateCenterABV$ABV)
abline(lm(StateCenterABV$ABV~StateCenterABV$Longitude),col="blue",lwd=2)
cor.test(StateCenterABV$Longitude, StateCenterABV$ABV)
```

# A Look at possible correlation between latitude and ABV
```{r LatLong1, ECHO=TRUE}
StateLatLong <- data.frame(matrix(ncol = 3, nrow = 50))
sa <- as.data.frame(state.abb)
names(sa)[names(sa) == "state.abb"] <- "State"
sc <- as.data.frame(state.center)
StateLatLong <- cbind(sa,sc)
names(StateLatLong)[names(StateLatLong) == "x"] <- "Longitude"
names(StateLatLong)[names(StateLatLong) == "y"] <- "Latitude"
StateLatLong$State <- as.character(as.factor(StateLatLong$State))
state_ABV$State <- as.character(as.factor(state_ABV$State))
StateLatLong <- rbind(StateLatLong, c("DC", -77.0369, 38.9072)) #Adds DC to the list
StateLatLong$Longitude <- as.numeric(as.character(StateLatLong$Longitude))
StateLatLong$Latitude <- as.numeric(as.character(StateLatLong$Latitude))
StateLatLong <- StateLatLong[order(StateLatLong$State),]
StateCenterABV <-data.frame(matrix(ncol = 4, nrow = 51))
#StateCenterABV <- data.frame(matrix(ncol=4, nrow=51)
StateCenterABV <- merge(StateLatLong, state_ABV, by=c('State'))
StateCenterABV

cor.test(StateCenterABV$Latitude, StateCenterABV$ABV)
lm(StateCenterABV$ABV~StateCenterABV$Latitude) #in the form Y=X
myfit=lm(StateCenterABV$ABV~StateCenterABV$Latitude) #in the form Y=X
myfit
plot(StateCenterABV$Latitude, StateCenterABV$ABV)
abline(lm(StateCenterABV$ABV~StateCenterABV$Latitude),col="blue",lwd=2)
```





TODO: CONCLUSION (Decide what to say in the conclusion and if we need to add any more code.)
