---
title: "BrewData Case Study 01 - Beers"
author: "Jim Park & Randall Hendrickson"
date: "October 6, 2017"
output: 
  html_document:
      keep_md: true
params:
  state: "TX"
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
if(!require("here")) cat("please install here package # devtools::install_github(\"krlmlr/here\")")
library(knitr)
library(here)
dataSetup_script <- here("scripts", "dataSetup.R")
analysis_script <- here("scripts", "analysis.R")
source(dataSetup_script)
#source(analysis_script)
```

#### Case Study 01

Beers

#### Questions

```{r setup_codebook2, echo=FALSE}
read_chunk(paste(analysis_script))
```

#### 1. How many breweries are present in each state?
```{r script_Analysis_Question1, echo=FALSE}
```

```{r echo=TRUE}
number_breweries_byState
```

#### 2. Merge beer data with the brewery data.
```{r script_Analysis_Question2, echo=FALSE}
```

```{r echo=TRUE}
# print first 6 observations
head(script_beer_world, 6)

# print last 6 observations
tail(script_beer_world, 6)
```

#### 3. Report the number of NS's in each column
```{r script_Analysis_Question3, echo=FALSE}
```

```{r echo=TRUE}
# report number of NAs
num_NAs
```

#### 4. Compute the median alcohol content and international bitterness unit for each state.
```{r script_Analysis_Question4, echo=TRUE}
```

#### 5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r script_Analysis_Question5, echo=FALSE}
```

```{r echo=TRUE}
# max ABV
script_state_max_ABV

# max ABV
script_state_max_IBU
```

#### 6. Summary statistics for the ABV variable.
```{r script_Analysis_Question6, echo=TRUE}
```

#### 7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
```{r script_Analysis_Question7, echo=TRUE}
```



```{r state_to_analyze}

cat(paste("This is the state to analyze:", params$state))

```


```{r child = './analysis.Rmd'}
```

```{r child = './customer.Rmd'}
```

## Reproducible Research

Include the session info, e.g. this document is produced with **knitr**. 
Here is the session info:


```{r session-info}
print(sessionInfo(), locale=FALSE)

```




