---
title: "MarkdownProject1"
output: md_document
---

## Number of Breweries by State
```{r, ECHO=TRUE}
devtools::install_github("krlmlr/here") #Commented out because it is only need to install once.
library(here)
# setwd("C:/Users/Jim/Google Drive/School/DoingDataScience/Proj1")
local_file1 <- here("data", "Breweries.csv")
local_file2 <- here("data", "Beers.csv")
Breweries <- read.csv(paste(local_file1),sep=",", header = TRUE)
head(Breweries, n=25)
library('plyr') # to call the count command
States <- (Breweries$State) #assigns variable to States
count(States) #Generates frequency table.  Number of Observations(breweries)/State
```

## Unit test for brewery names repeated
```{r UnitTest_BreweriesByState, ECHO=TRUE}
Breweries <- read.csv(paste(local_file1),sep=",", header = TRUE)
head(Breweries, n=25)
length(unique(States)) #counts unique categories in major_count. 51 States represented
length(unique(Breweries$Name))
n_occur <- data.frame(table(Breweries$Name))# gives you a data frame with a list of ids and the number of times they occurred.
n_occur[n_occur$Freq > 1,]
```

##Merged data sets Print first 6 and last 6
```{r Q2MergedData, ECHO=TRUE}

Beers <- read.csv(paste(local_file2),sep=",", header = TRUE)
mergedData <- merge(Breweries, Beers, by.x=c('Brew_ID'), by.y=c('Brewery_id'))
names(mergedData)[names(mergedData) == "Name.x"] <- "Brewery_Name"
names(mergedData)[names(mergedData) == "Name.y"] <- "Beer_Name"
head(mergedData,n=6)
tail(mergedData, n=6)
```

## Report the number of NA's in each column
```{r Q3NumberNAsInEachCol, ECHO=TRUE}
summary(mergedData)
```
IBU has 1005 NA's and ABV has 62 NA's.




## FUNCTION BUILDING

# Create list of States by alphabetical oarder for user to copy into function.
```{r StateListAphabetical, ECHO=TRUE}
library('plyr')
ListOfStates <- sort(unique(mergedData$State))
ListOfStates <- data.frame(lapply(ListOfStates, as.character), stringsAsFactors=FALSE)
as.character(ListOfStates)
ListOfStates <- gsub(" ", "", ListOfStates, fixed = TRUE) #Removes extra spaces that area not needed.
```
## Compute the median Alcohol By Volume (ABV) and international bitterness unit (IBU) for each state.  Plot a bar chart to compare.
```{r StateABVIBUTable, ECHO=TRUE}
state_ABV <- aggregate(mergedData["ABV"], by=mergedData[c("State")], FUN=median, na.rm=TRUE)
State_IBU <- aggregate(mergedData["IBU"], by=mergedData[c("State")], FUN=median, na.rm=TRUE)
ABV_IBU_Output <- merge(state_ABV, State_IBU, by=c('State'))
ABV_IBU_Output
```
## Compute the number of beers in the state selected. Input State Desired.
```{r BeersInState, ECHO=TRUE}
StateSelected <- c('MI') #Copy and paste desited state inbetween the quotes here from list of States by alphabetical order.
StateBeers <- mergedData[grep(StateSelected, mergedData$State),]
BeersInStateOutput <- paste('There are', length(ListOfStates), ' Unique states in this data set and', nrow(StateBeers), 'unique beers in the state of', StateSelected)
BeersInStateOutput
StateBeers
```

## Input State desired.  Plots ABV and Median IBU with state highlighted
```{r ABVIBUCharts, ECHO=TRUE}
#FUNCTION START - UNCOMMENT LATER
#StateOutput <- function (StateSelected, ListOfStates){
StateSelected <- c('MI') #Copy and paste desited state inbetween the quotes here from list of States by alphabetical order.
library('plyr')
ListOfStates <- sort(unique(mergedData$State))
ListOfStates #Just a list of the states in alphabetical order assuming it may be needed by the application user.
StateQty <- data.frame(matrix(ncol = 2, nrow = 51)) #Initiates data frame for loop
ColorOfPlot <- 0 #Initiates vector for loop to be used to color code bar chart.
library(data.table)
setnames(StateQty, old = c('X1','X2'), new = c('State','Number')) #Changes the names of columns.

#  Loop below builds a data frame with state, number of beers in each state, and color scheme for bar chart
for(i in 1:51){
StateQty$State[i] <- as.character(ListOfStates[i]) #Adds each state to the vector
StateBeers <- mergedData[grep(ListOfStates[i], mergedData$State),]
StateQty$Number[i] <- nrow(StateBeers)
ColorOfPlot[i] <- 'Red' #Initializes all to the color red.
}

StateQty <- StateQty[order(-StateQty$Number),]#Orders from high to low based on number in brewery
#StateQty
ColorOfPlot[which(grepl(StateSelected, StateQty$State))] <- 'blue' #Changes the color of the state of interest to stand out in plotting.  Feeds into fill command in plot that follows.
# Plots a histogram in decreasing order of beers per state with the state of interest highlighted
library(ggplot2)
ggplot(data=StateQty, aes(x=reorder(State, -Number), y=Number)) + geom_bar(stat = "identity", fill=ColorOfPlot) +  theme(axis.text.x=element_text(angle=90,vjust=0.5)) + xlab("State") + ylab("Number of Beers") + ggtitle("Number of Different Beers by State")

# Loop to construct the color scheme for fill option in plotting median IBU by state
ColorOfPlotIBU <- 0
for(i in 1:51){
ColorOfPlotIBU[i] <- 'Red'
}
State_IBU <- State_IBU[order(-State_IBU$IBU),]#Orders from high to low based on IBU
State_IBU[is.na(State_IBU)] <- 0
ColorOfPlotIBU[which(grepl(StateSelected, State_IBU$State))] <- 'blue' 
ggplot(data=State_IBU, aes(x=reorder(State, -IBU), y=IBU)) + geom_bar(stat = "identity", fill=ColorOfPlotIBU) +  theme(axis.text.x=element_text(angle=90,vjust=0.5)) + xlab("State") + ylab("IBU") + ggtitle("IBU Median by State") + labs(caption = "(*SD did not have any reported IBU values)")

# }  function end. uncomment later.
```



```{r, ECHO=TRUE}
# Function Setup and inputs
#StateReport <- function(StateOfInterest, Filename) {
#FileToRead <- paste(getwd(), '/', Filename, sep="")
#FileToRead
#SleepData <- read.csv(FileToRead)

# 3a. Create Objects
#MedianAge <- median(SleepData$Age, na.rm=TRUE) #Median Age Object
#MinDurationSleep <- min(SleepData$Duration, na.rm=TRUE) #Min Sleep Object
#MaxDurationSleep <- max(SleepData$Duration, na.rm=TRUE) # Max Sleep object
#MeanRSES <- mean(SleepData$RSES, na.rm=TRUE) # Mean RSES Object
#StdDevRSES <- sd(SleepData$RSES, na.rm=TRUE) # Standard Deviation of RSES Object
#SumObjects <- c(MedianAge, MinDurationSleep, MaxDurationSleep, MeanRSES, StdDevRSES)
#SumObjects #Creates one vector with all of the values.

#3b. Create a data.frame
#report <- data.frame(MedianAge, MeanRSES/5, StdDevRSES/5, MaxDurationSleep - MinDurationSleep) # creates data frame with the requested data for question 3b.
#report

#3c. Change the column names
#cnames <- c("MedianAge", "SelfEsteem", "SE_SD", "DurationRange")
#colnames(report) <- cnames
#report

#3d. Round the report to at most 2 digits
#report$SelfEsteem <- round(report$SelfEsteem, digits = 2)
#report$SE_SD <- round(report$SE_SD, digits = 2)
#report
#}
```

3e Run Function

```{r, RunFunction, ECHO=TRUE}
#setwd("c:/Users/Jim/Google Drive/school/DoingD~1/Week3")
#Filename <- c('sleep_data_01.csv')
#SummaryReport(Filename)
```


